---
title: "Data Science Dashboard"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: fill
    # vertical_layout: fill
#runtime: shiny
---

```{r setup, include=FALSE}
library(dplyr)
library(flexdashboard)
library(igraph)
library(readxl)
library(sf)
library(sfnetworks)
library(tidygraph)
library(tidyverse)
library(tmap)
library(geosphere)

tmap_mode("view")
```

Trains in London {data-orientation=rows}
=====================================
```{r, include=FALSE}
Stations <- st_read("data/tfl_stations_new.json")
Lines <- st_read("data/tfl_lines_new.json")
Zones <- st_read("data/tfl_zones.json") %>%
  rename("Name" = "name")

colors <- c(
  "#B26300", # Bakerloo
  "#DC241F", # Central
  "#FFD329", # Circle
  "#9364CC", # Crossrail
  "#e1eb5b", # Crossrail 2
  "#007D32", # District
  "#00AFAD", # DLR
  "#E0A9BE", # Hammersmith & City
  "#A1A5A7", # Jubliee
  "#9B0058", # Metropolitan
  "#000000", # Northern
  "#EF7B10", # Overground
  "#0019A8", # Piccadilly
  "#0098D8", # Victoria
  "#93CEBA" # Waterloo & City
)

tube_lines <- c("Bakerloo", "Central", "Circle", "District", "Hammersmith", "Jubilee", "Metropolitan", "Northern", "Piccadilly", "Victoria", "Waterloo")

num_station <- nrow(Stations)

rail_len <- st_length(Lines) %>%
  sum()

num_lines <- 15
```
Row
-------------------------------------

### Total number of train stations by 2030s

```{r}
valueBox(num_station, icon = "fa-train")
```

### Number of operated lines (Overground and DLR count as one single line)
```{r}
valueBox(num_lines, icon = "fa-subway")
```

### Total track length

```{r}
valueBox(sprintf("%d km", round(rail_len / 1000)))
```

Row
-------------------------------------

### Tube System Map including Overground, DLR and Crossrail
```{r}
Stations2 <-  Stations %>%
  rename("Number of Lines" = "Number.of.Lines", "Station ID" = "Station.ID", "Line(s)" = "Lines")
tm_shape(Zones)+
  tm_polygons(col="Name", alpha=.2, legend.show=FALSE)+
  tm_shape(Lines) +
  tm_lines(col="Line", scale=3, palette = colors) +
  tm_shape(Stations2) +
  tm_dots(scale=1.5)
```

Trains at Night {data-orientation=rows}
===============
```{r, include=FALSE}
Stations <- st_read("data/tfl_stations_night.json")
Lines <- st_read("data/tfl_lines_night.json")
colors_night <- c(
  "#DC241F", # Central
  "#A1A5A7", # Jubliee
  "#000000", # Northern
  "#EF7B10", # Overground
  "#0019A8", # Piccadilly
  "#0098D8" # Victoria
)

num_station <- nrow(Stations)

rail_len <- st_length(Lines) %>%
  sum()

num_lines <- 6

```
Row
-------------------------------------
### Total number of train stations opened at night

```{r}
valueBox(num_station, icon = "fa-train")
```

### Number of operated lines during the night (Overground and DLR count as one single line)
```{r}
valueBox(num_lines, icon = "fa-subway")
```

### Total track length that is being used during the night

```{r}
valueBox(sprintf("%d km", round(rail_len / 1000)))
```

Row
-------------------------------------

### London Night Tube map including Overground
```{r}
tm_shape(Zones)+
  tm_polygons(col="Name", alpha=.2, legend.show=FALSE)+
  tm_shape(Lines) +
  tm_lines(col="Line", palette=colors_night, scale=3) +
  tm_shape(Stations) +
  tm_dots(scale=1.5)
```



National Rail {data-orientation=rows}
===============

```{r, include=FALSE}
nr_stations <- st_read("data/nr_stations.json")
nr_lines <- st_read("data/nr_lines.json")

num_stations <- nrow(nr_stations)
rail_len <- st_length(nr_lines) %>%
  sum()

```

Row
-------------------------------------
### Total number of train stations operated by National Rail

```{r}
valueBox(num_stations, icon = "fa-train")
```

### Total track length of the National Rail

```{r}
valueBox(sprintf("%d km", round(rail_len / 1000)))
```

Row
-------------------------------------

### Map of the National Rail in Great Britain

```{r}
tm_shape(nr_lines) +
  tm_lines() +
  tm_shape(nr_stations) +
  tm_dots()
```

Underpasses
================

```{r, include=FALSE}
# Import London map of boroughs
boroughs <- st_read("data/gis/London_Borough_Excluding_MHW.shp")
boroughs <- select(boroughs, NAME, GSS_CODE, HECTARES, NONLD_AREA, geometry)

height_restrictions <- read_excel("data/london_height-restrictions.xlsx")
# Extract simplified underpass data before conversion
underpasses_road_number <- select(height_restrictions, `Road number`)
# Extract categories from text and convert to double:
height_restrictions <- mutate(height_restrictions, `height_temp` = str_replace_all(`Height restriction (m)`, c("Between " = "", "Up to 3.0" = "0")))
height_restrictions <- mutate(height_restrictions, height = as.double(str_sub(`height_temp`, 1, 3)))
height_restrictions <- st_as_sf(height_restrictions, coords = c("Lng", "Lat"), crs = 4326) 

# Read data from OSM CSV export
osm_underpasses <- read.csv2("data/london_underpasses.csv", encoding = "UTF-8", sep = "\t", na.strings = c(""))
# data without a maximum clearance is not useful and mostly due to complex geometries
osm_underpasses <- filter(osm_underpasses, !is.na(maxheight))
osm_underpasses <- st_as_sf(osm_underpasses, coords = c("X.lon", "X.lat"), crs = 4236)
osm_underpasses <- st_transform(osm_underpasses, crs = 3875)

# Convert text strings to numbers
# Extension: convert imperial measurements to metres
osm_underpasses <- mutate(osm_underpasses, height = as.numeric(maxheight))
```

Column {data-width=350}
-----------------------------------------------------------------------

### Underpasses based on National Statistics and Ordnance Survey

```{r}
tm_shape(boroughs) + tm_polygons() + tm_shape(height_restrictions) + tm_dots(col = "height")
```

> Contains National Statistics data © Crown copyright and database right [2015]
> Contains Ordnance Survey data © Crown copyright and database right [2015]

### Underpasses by borough (OS)

```{r}
ggplot(height_restrictions, mapping=aes(x = `Borough`, fill=factor(height))) + geom_bar() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0.0, vjust = 0.2)) +
  scale_y_continuous("Number of underpasses") +
  guides(fill = guide_legend(title = "minimum\nclearance"))
# X axis text orientation: https://ggplot2-book.org/polishing.html#theme-axis
```

Column {data-width=350}
-----------------------------------------------------------------------

### Underpasses based on OpenStreetMap

```{r}
tm_shape(boroughs) + tm_polygons() + tm_shape(osm_underpasses) + tm_dots(col = "height")
```

> © [OpenStreetMap](https://osm.org/copyright) contributors

### Underpasses by road number (OS)

```{r}
# Data queried via https://overpass-api.de/query_form.html
# [out:csv("name","highway","maxheight",::lat,::lon)]
# [timeout:180]
# ;
# (
#   way
#     ["maxheight"]
#     ["highway"]
#     (area:3600175342);
# );
# out center;
# >;
# out;
underpasses_road_number <- filter(underpasses_road_number, !is.na(`Road number`))
underpasses_road_number <- group_by(underpasses_road_number, `Road number`)
underpasses_road_number <- summarise(underpasses_road_number, number = n())
underpasses_road_number <- arrange(underpasses_road_number, desc(number))
knitr::kable(underpasses_road_number)
```

Accidents
==========
```{r, include = FALSE}
accidents <- read.csv2("data/london_2019-road-accidents.csv", encoding = "UTF-8", sep = ",")
```

Data source <https://roads.data.tfl.gov.uk/AccidentStats/Dev/2019-gla-data-extract-vehicle.csv>

Column {data-width=350}
-------------------------------------------------------------------------------------------

### Accidents per borough
```{r}
boroughaccidents <- group_by(accidents, Borough)
boroughaccidents <- summarise(boroughaccidents, number = n())
boroughaccidents <- mutate(boroughaccidents, Borough = str_replace(Borough, "-UPON-THAMES", " UPON THAMES"))
boroughaccidents <- mutate(boroughaccidents, Borough = str_replace(Borough, " & ", " AND "))
boroughaccidents <- rename(boroughaccidents, BName = Borough)
ba_spatial <- left_join(mutate(boroughs, BName = str_to_upper(NAME)), boroughaccidents)
ba_spatial <- select(ba_spatial, NAME, GSS_CODE, HECTARES, NONLD_AREA, number, geometry)
ba_spatial <- rename(ba_spatial, Accidents = number)

# Coloring doc: https://geocompr.github.io/post/2019/tmap-color-scales/#pretty
tm_shape(ba_spatial) + tm_polygons(col = "Accidents") + tm_legend(legend.outside = TRUE)
```

### Percentage of female drivers in accidents
```{r, size=200}
driver_sex <- group_by(accidents, Driver.Sex)
driver_sex <- summarise(driver_sex, n = n())
female_drivers <- filter(driver_sex, Driver.Sex == "2 FEMALE")[2]
male_drivers <- filter(driver_sex, Driver.Sex == "1 MALE")[2]
unidentified_drivers <- filter(driver_sex, Driver.Sex == "3 NOT TRACED")[2]
valueBox(round(female_drivers / (female_drivers + male_drivers + unidentified_drivers) * 100, digits = 0), icon = "fa-female")
```

### Median driver age
```{r}
driver_age <- summarise(filter(accidents, Driver.Age != 0), age = median(Driver.Age))
valueBox(driver_age)
```


Column {data-width=350}
-------------------------------------------------------------------------------------------

### Vehicle manoeuvres
```{r}
manoeuvres <- transmute(accidents, manoeuvre = str_sub(Vehicle.Manoeuvres, 4), skid = str_sub(Vehicle.Skidding, 3))
manoeuvres <- group_by(manoeuvres, manoeuvre, skid)
manoeuvres <- summarise(manoeuvres, number = n())
knitr::kable(manoeuvres)
```


<!-- Bus Line App -->
<!-- =====================================  -->

<!-- Column {.sidebar}  -->
<!-- ----------------------------------------------------------------------- -->

<!-- ```{r} -->
<!-- data_day <- read.csv("data/all_bus_lines.csv") %>% -->
<!--   mutate(switch = "day") %>% -->
<!--   select(-X.1, -X) -->

<!-- data_night <- read.csv("data/all_bus_lines_night.csv") %>% -->
<!--   mutate(switch = "night") %>% -->
<!--   select(-X) -->

<!-- data <- rbind(data_day, data_night) -->

<!-- #choices <- as.character(data$lineId) -->

<!-- #selectInput("lineId",  -->
<!-- #            "Select Line: ",  -->
<!-- #            unique(choices), -->
<!-- #            selected = "1", -->
<!-- #            multiple = FALSE) -->

<!-- ``` -->
<!-- Row  -->
<!-- ----------------------------------------------------------------------- -->

<!-- ### From -->

<!-- ```{r} -->
<!-- #renderValueBox({ -->

<!-- #start_point <- data %>% -->
<!-- #  filter(lineId == input$lineId) %>% -->
<!-- #  select(name) -->

<!-- #valueBox(start_point$name[1],  -->
<!-- #         icon = "fa-arrow-circle-down") -->
<!-- #}) -->

<!-- ``` -->

<!-- ### To -->

<!-- ```{r} -->
<!-- #renderValueBox({ -->

<!-- #end_point <- data %>% -->
<!-- #  filter(lineId == input$lineId) -->
<!-- #  index <- end_point %>% -->
<!-- #  nrow() -->

<!-- #valueBox(end_point$name[index],  -->
<!-- #         icon = "fa-arrow-circle-up") -->

<!-- #}) -->

<!-- ``` -->

<!-- ### Number of stops -->

<!-- ```{r} -->
<!-- #renderValueBox({ -->

<!-- #number_stops <- data %>% -->
<!-- #  filter(lineId == input$lineId) %>% -->
<!-- #  nrow() -->

<!-- #valueBox(number_stops, -->
<!-- #          icon = "fa-bus-alt")     -->
<!-- #}) -->

<!-- ``` -->

<!-- ### Distance -->

<!-- ```{r} -->
<!-- #renderValueBox({ -->

<!-- #distance <- data %>% -->
<!-- #  filter(lineId == input$lineId) %>% -->
<!-- #  st_as_sf(coords = c("Location_Easting", "Location_Northing"), crs = 27700) %>% -->
<!-- #  group_by(lineId) %>% -->
<!-- #  summarise(do_union = FALSE) %>% -->
<!-- #  st_cast("LINESTRING") %>% -->
<!-- #  st_length() -->

<!-- #distance <- (distance/1000) %>% -->
<!-- #  round(digits = 2) -->

<!-- #valueBox(paste(distance, "km"), -->
<!-- #          icon = "fa-road") -->
<!-- #}) -->

<!-- ``` -->

<!-- ### Time -->

<!-- ```{r} -->
<!-- #renderValueBox({ -->

<!-- #data_dist <- data %>% -->
<!-- #  filter(lineId == input$lineId) -->

<!-- #distance <- data_dist %>% -->
<!-- #  st_as_sf(coords = c("Location_Easting", "Location_Northing"), crs = 27700) %>% -->
<!-- #  group_by(lineId) %>% -->
<!-- #  summarise(do_union = FALSE) %>% -->
<!-- #  st_cast("LINESTRING") %>% -->
<!-- #  st_length() -->

<!-- #avg_speed <- (data_dist$Average_Speed[1] / 1.609) -->

<!-- #time <- (((as.numeric(distance)/1000)*60)/(avg_speed)) %>% -->
<!-- #  round(digits = 0) -->

<!-- #valueBox(paste(time, "min"), -->
<!-- #          icon = "fa-clock") -->
<!-- #}) -->

<!-- ``` -->
<!-- Column {data-width=650} -->
<!-- ----------------------------------------------------------------------- -->

<!-- ### Route -->

<!-- ```{r} -->
<!-- #bus_line <- reactive({ -->
<!-- #  data %>% -->
<!-- #    filter(lineId == input$lineId) %>% -->
<!-- #    st_as_sf(coords = c("Location_Easting", "Location_Northing"), crs = 27700) -->
<!-- # }) -->

<!-- #tmap_mode("view") -->

<!-- #renderTmap({ -->

<!-- #net <- bus_line() %>%   -->
<!-- #  group_by(lineId) %>% -->
<!-- #  summarise(do_union = FALSE) %>% -->
<!-- #  st_cast("LINESTRING") %>% -->
<!-- #  as_sfnetwork(directed = TRUE)   -->

<!-- #tm_shape(st_as_sf(net, "edges")) +  -->
<!-- #  tm_lines(id = c("Line" = "lineId"), -->
<!--            scale = 2) + -->
<!-- #  tm_shape(bus_line()) +  -->
<!-- #  tm_dots(col = "#dc241f", -->
<!-- #          size = 0.04,  -->
<!-- #          id = "name") +  -->
<!-- #    tm_shape(st_as_sf(net, "nodes")) +  -->
<!-- #    tm_dots(col = "#0019a8",  -->
<!-- #            size = 0.1, -->
<!-- #            id = "name") -->
<!-- #}) -->

<!-- ``` -->
<!-- Network Analysis - Nightbuses -->
<!-- =====================================  -->

<!-- Column {data-width=650} -->
<!-- ----------------------------------------------------------------------- -->

<!-- ### Betweenness centrality -->

<!-- ```{r} -->

<!-- # Simplified night bus network -->

<!-- bus_lines_mp <- read.csv("data/all_bus_lines_night.csv") %>% -->
<!--   filter(!is.na(Location_Easting)) %>% -->
<!--   group_by(parentId) %>% -->
<!--   mutate(Location_Easting = mean(Location_Easting)) %>% -->
<!--   mutate(Location_Northing = mean(Location_Northing)) -->

<!-- # Simplified bus stops -->
<!-- stop_mp <- bus_lines_mp %>% -->
<!--   select(parentId, stationId, name, Location_Easting, Location_Northing) %>% -->
<!--   unique() -->

<!-- # Create edges and nodes for sfnetworks -->

<!-- nodes <- bus_lines_mp %>% -->
<!--   select(X, lineId,parentId, Location_Easting, Location_Northing, name) %>% -->
<!--   rename("nodeID" = X) %>% -->
<!--   mutate(start = "start", end = "end") %>% -->
<!--   gather(start_end, end, start:end) -->

<!-- nodes_start <- nodes %>% -->
<!--   filter(start_end == "start") %>% -->
<!--   group_by(lineId) %>%  -->
<!--   mutate(Index_Group = row_number()) %>% -->
<!--   filter(Index_Group != max(n())) %>% -->
<!--   ungroup() -->

<!-- nodes_end <- nodes %>% -->
<!--   filter(start_end == "end") %>% -->
<!--   group_by(lineId) %>%  -->
<!--   mutate(Index_Group = row_number()) %>% -->
<!--   filter(Index_Group != 1) %>% -->
<!--   ungroup() %>% -->
<!--   mutate(nodeID = nodeID + 10000)   -->

<!-- edges <- cbind(D1 = nodes_start, D2 = nodes_end) %>% -->
<!--   rename(c(D1.lineId, D1.Location_Easting, D1.Location_Northing, D2.Location_Easting, D2.Location_Northing), -->
<!--          "lineId" = D1.lineId, -->
<!--          "lat_start" = D1.Location_Easting,  -->
<!--          "lon_start" = D1.Location_Northing,  -->
<!--          "lat_end" = D2.Location_Easting,  -->
<!--          "lon_end" = D2.Location_Northing) %>% -->
<!--   select(lineId, D1.name, lat_start, lon_start, lat_end, lon_end) %>% -->
<!--   unite(start_point, lat_start:lon_start, sep = "__") %>% -->
<!--   unite(end_point, lat_end:lon_end, sep = "__") %>% -->
<!--   mutate(Index = row_number()) %>%  -->
<!--   gather(D1.name, point, start_point:end_point) %>% -->
<!--   arrange(Index) %>% -->
<!--   separate(point, c("lat", "lon"), sep = "__") %>% -->
<!--   st_as_sf(coords = c("lat", "lon"), crs = 27700) %>% -->
<!--   group_by(Index) %>% -->
<!--   summarise(do_union = FALSE) %>% -->
<!--   st_cast("LINESTRING") -->

<!-- edges <- edges %>% -->
<!--   mutate(edgeID = c(1:n())) -->

<!-- nodes <- edges %>% -->
<!--   st_coordinates() %>% -->
<!--   as_tibble() %>% -->
<!--   rename(edgeID = L1) %>% -->
<!--   group_by(edgeID) %>% -->
<!--   slice(c(1, n())) %>% -->
<!--   ungroup() %>% -->
<!--   mutate(start_end = rep(c('start', 'end'), times = n()/2)) %>% -->
<!--   mutate(xy = paste(.$X, .$Y)) %>%  -->
<!--   mutate(nodeID = group_indices(., factor(xy, levels = unique(xy)))) %>% -->
<!--   select(-xy) -->

<!-- source_nodes <- nodes %>% -->
<!--   filter(start_end == "start") %>% -->
<!--   pull(nodeID) -->

<!-- target_nodes <- nodes %>% -->
<!--   filter(start_end == "end") %>% -->
<!--   pull(nodeID) -->

<!-- edges = edges %>% -->
<!--   mutate(from = source_nodes, to = target_nodes) -->

<!-- nodes <- nodes %>% -->
<!--   distinct(nodeID, .keep_all = TRUE) %>% -->
<!--   st_as_sf(coords = c('X', 'Y'), crs = 27700) %>% -->
<!--   select(nodeID, geometry) -->
<!--   #st_set_crs(st_crs(edges)) -->

<!-- network <- sfnetwork(nodes, edges, directed = TRUE) %>% -->
<!--   activate(edges) %>% -->
<!--   mutate(length = st_length(geometry)) -->

<!-- network <- network %>%   -->
<!--   activate(nodes) %>% -->
<!--   mutate(degree = centrality_degree()) %>% -->
<!--   mutate(betweenness = centrality_betweenness(weights = as.numeric(length) + 0.0001)) %>% -->
<!--   mutate(eigen = centrality_eigen()) %>% -->
<!--   activate(edges) %>% -->
<!--   mutate(betweenness = centrality_edge_betweenness(weights = as.numeric(length) + 0.0001)) -->

<!-- tm_shape(network %>% activate(edges) %>% as_tibble() %>% st_as_sf()) +  -->
<!--   tm_lines() +  -->
<!--   tm_shape(network %>% activate(nodes) %>% as_tibble() %>% st_as_sf()) + -->
<!--   tm_dots(col = "betweenness", size = "betweenness", palette = "cividis") -->

<!-- ``` -->


<!-- Column {data-width=650} -->
<!-- ----------------------------------------------------------------------- -->

<!-- ### Betweenness centrality -->

<!-- ```{r} -->

<!-- tm_shape(network %>% activate(edges) %>% as_tibble() %>% st_as_sf()) +  -->
<!--   tm_lines(col = "betweenness") +  -->
<!--   tm_shape(network %>% activate(nodes) %>% as_tibble() %>% st_as_sf()) + -->
<!--   tm_dots(col = "eigen", size = "degree", palette = "cividis") -->


<!-- ``` -->
