---
title: "Data Science Dashboard"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: fill
    # vertical_layout: fill
---

```{r setup, include=FALSE}
library(dplyr)
library(flexdashboard)
library(igraph)
library(readxl)
library(sf)
# library(sfnetworks)
library(tidygraph)
library(tidyverse)
library(tmap)
library(geosphere)

tmap_mode("view")
```

Underground
=====================================

Row
-------------------------------------
### Tube System Map including DLR

```{r, fig.width=10, fig.height=7}
lines <- read_csv("data/london_lines.csv")
connections = read_csv("data/london_connections.csv", col_types = cols(station1=col_integer(), station2=col_integer())) %>%
  rename(from=station1, to=station2)
stations <- read_csv("data/london_stations.csv", col_types = cols(id = col_integer()))

nodes <- stations %>%
  st_as_sf(coords=c("longitude", "latitude")) %>%
  select("Station" = "name", "Number of Lines"="total_lines", "Railway"="rail", "position"="geometry") %>%
  mutate(Railway = ifelse(Railway == 1, "Yes", "No"))

edges <- connections %>%
  inner_join(stations, c("from" = "id")) %>%
  inner_join(stations, c("to" = "id")) %>%
  mutate(wkt=sprintf(
    "LINESTRING(%f %f, %f %f)", longitude.x, latitude.x, longitude.y, latitude.y
  )) %>%
  inner_join(lines, c("line" = "line")) %>%
  rowwise() %>%
  mutate(distance = distm( x = c(longitude.x, latitude.x), y = c(longitude.y, latitude.y), fun=distHaversine))


edges <- st_as_sf(edges, wkt="wkt", dim ="XY") %>%
  mutate(distance = sprintf(
    "%.0f m", distance
  )) %>%
  select("Distance" = "distance", "wkt" = "wkt", "Line" = "name", "Station 1" = "name.x", "Station 2" = "name.y")

tm_shape(edges)+
  tm_lines(col="Line", scale=3)+
  tm_shape(nodes) +
  tm_dots(text="Station", size = 0.02) 

e <- edges %>%
  mutate(dist = as.numeric(strsplit(Distance, " ")[[1]][1])) %>%
  arrange(desc(dist))
st_geometry(e) <- NULL
avg = group_by(e, Line) %>%
  summarise(avg_dist = mean(dist), num = n()) %>%
  arrange(desc(avg_dist)) %>%
  mutate(avg_dist = sprintf(
    "%.0f m", round(avg_dist
  ))) %>%
  rename("Average distance" = avg_dist, "Number of stations" = num)
e <- e[, -5]
```
Row
-------------------------------------
### Distance between two Tube Stations

```{r}
knitr::kable(e)
```

### Average distance between two stations by Tube Line
```{r}
knitr::kable(avg)
```

Tube Network Analysis
===============

### Total number of Tube stations including DLR

```{r}
number <- nrow(nodes)
valueBox(number, icon = "fa-subway")
```

Underpasses
================

```{r, include=FALSE}
# Import London map of boroughs
boroughs <- st_read("data/gis/London_Borough_Excluding_MHW.shp")
boroughs <- select(boroughs, NAME, GSS_CODE, HECTARES, NONLD_AREA, geometry)

height_restrictions <- read_excel("data/london_height-restrictions.xlsx")
# Extract simplified underpass data before conversion
underpasses_road_number <- select(height_restrictions, `Road number`)
# Extract categories from text and convert to double:
height_restrictions <- mutate(height_restrictions, `height_temp` = str_replace_all(`Height restriction (m)`, c("Between " = "", "Up to 3.0" = "0")))
height_restrictions <- mutate(height_restrictions, height = as.double(str_sub(`height_temp`, 1, 3)))
height_restrictions <- st_as_sf(height_restrictions, coords = c("Lng", "Lat"), crs = 4326) 

# Read data from OSM CSV export
osm_underpasses <- read.csv2("data/london_underpasses.csv", encoding = "UTF-8", sep = "\t", na.strings = c(""))
# data without a maximum clearance is not useful and mostly due to complex geometries
osm_underpasses <- filter(osm_underpasses, !is.na(maxheight))
osm_underpasses <- st_as_sf(osm_underpasses, coords = c("X.lon", "X.lat"), crs = 4236)
osm_underpasses <- st_transform(osm_underpasses, crs = 3875)

# Convert text strings to numbers
# Extension: convert imperial measurements to metres
osm_underpasses <- mutate(osm_underpasses, height = as.numeric(maxheight))
```

Column {data-width=350}
-----------------------------------------------------------------------

### Underpasses based on National Statistics and Ordnance Survey

```{r}
tm_shape(boroughs) + tm_polygons() + tm_shape(height_restrictions) + tm_dots(col = "height")
```

> Contains National Statistics data © Crown copyright and database right [2015]
> Contains Ordnance Survey data © Crown copyright and database right [2015]

### Underpasses by borough (OS)

```{r}
ggplot(height_restrictions, mapping=aes(x = `Borough`, fill=factor(height))) + geom_bar() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0.0, vjust = 0.2)) +
  scale_y_continuous("Number of underpasses") +
  guides(fill = guide_legend(title = "minimum\nclearance"))
# X axis text orientation: https://ggplot2-book.org/polishing.html#theme-axis
```

Column {data-width=350}
-----------------------------------------------------------------------

### Underpasses based on OpenStreetMap

```{r}
tm_shape(boroughs) + tm_polygons() + tm_shape(osm_underpasses) + tm_dots(col = "height")
```

> © [OpenStreetMap](https://osm.org/copyright) contributors

### Underpasses by road number (OS)

```{r}
# Data queried via https://overpass-api.de/query_form.html
# [out:csv("name","highway","maxheight",::lat,::lon)]
# [timeout:180]
# ;
# (
#   way
#     ["maxheight"]
#     ["highway"]
#     (area:3600175342);
# );
# out center;
# >;
# out;
underpasses_road_number <- filter(underpasses_road_number, !is.na(`Road number`))
underpasses_road_number <- group_by(underpasses_road_number, `Road number`)
underpasses_road_number <- summarise(underpasses_road_number, number = n())
underpasses_road_number <- arrange(underpasses_road_number, desc(number))
knitr::kable(underpasses_road_number)
```

Accidents
==========
```{r, include = FALSE}
accidents <- read.csv2("data/london_2019-road-accidents.csv", encoding = "UTF-8", sep = ",")
```

Data source <https://roads.data.tfl.gov.uk/AccidentStats/Dev/2019-gla-data-extract-vehicle.csv>

Column {data-width=350}
-------------------------------------------------------------------------------------------

### Accidents per borough
```{r}
boroughaccidents <- group_by(accidents, Borough)
boroughaccidents <- summarise(boroughaccidents, number = n())
boroughaccidents <- mutate(boroughaccidents, Borough = str_replace(Borough, "-UPON-THAMES", " UPON THAMES"))
boroughaccidents <- mutate(boroughaccidents, Borough = str_replace(Borough, " & ", " AND "))
boroughaccidents <- rename(boroughaccidents, BName = Borough)
ba_spatial <- left_join(mutate(boroughs, BName = str_to_upper(NAME)), boroughaccidents)
ba_spatial <- select(ba_spatial, NAME, GSS_CODE, HECTARES, NONLD_AREA, number, geometry)
ba_spatial <- rename(ba_spatial, Accidents = number)

# Coloring doc: https://geocompr.github.io/post/2019/tmap-color-scales/#pretty
tm_shape(ba_spatial) + tm_polygons(col = "Accidents") + tm_legend(legend.outside = TRUE)
```

### Percentage of female drivers in accidents
```{r, size=200}
driver_sex <- group_by(accidents, Driver.Sex)
driver_sex <- summarise(driver_sex, n = n())
female_drivers <- filter(driver_sex, Driver.Sex == "2 FEMALE")[2]
male_drivers <- filter(driver_sex, Driver.Sex == "1 MALE")[2]
unidentified_drivers <- filter(driver_sex, Driver.Sex == "3 NOT TRACED")[2]
valueBox(round(female_drivers / (female_drivers + male_drivers + unidentified_drivers) * 100, digits = 0), icon = "fa-female")
```

### Median driver age
```{r}
driver_age <- summarise(filter(accidents, Driver.Age != 0), age = median(Driver.Age))
valueBox(driver_age)
```


Column {data-width=350}
-------------------------------------------------------------------------------------------

### Vehicle manoeuvres
```{r}
manoeuvres <- transmute(accidents, manoeuvre = str_sub(Vehicle.Manoeuvres, 4), skid = str_sub(Vehicle.Skidding, 3))
manoeuvres <- group_by(manoeuvres, manoeuvre, skid)
manoeuvres <- summarise(manoeuvres, number = n())
knitr::kable(manoeuvres)
```

Bus
=====================================