---
title: "Data Science Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(ggplot2)
library(forcats)
library(httr)
library(jsonlite)
library(purrr)
library(readxl)
library(openxlsx)
library(sf)
library(sfnetworks)
library(tidygraph)
library(tmap)
library(plotly)

data_day <- read.csv("data/all_bus_lines.csv") %>%
  mutate(switch = "day") %>%
  select(-X.1, -X)
data_night <- read.csv("data/all_bus_lines_night.csv") %>%
  mutate(switch = "night") %>%
  select(-X)

```
Bus Line App
===================================== 

COlumn {.sidebar} 
-----------------------------------------------------------------------

```{r}
data <- rbind(data_day, data_night)

choices <- as.character(data$lineId)

selectInput("lineId", 
          "Select Line: ", 
          unique(choices),
          selected = "1",
          multiple = FALSE)

```

Row 
-----------------------------------------------------------------------

### From

```{r}


renderValueBox({

  start_point <- data %>%
    filter(lineId == input$lineId) %>%
    select(name)

valueBox(start_point$name[1], 
          icon = "fa-arrow-circle-down")

})


```

### To


```{r}

renderValueBox({
  
  end_point <- data %>%
    filter(lineId == input$lineId)
    index <- end_point %>%
    nrow()
  
  valueBox(end_point$name[index], 
           icon = "fa-arrow-circle-up")
  
})

```

### Number of stops

```{r}


renderValueBox({
 
number_stops <- data %>%
  filter(lineId == input$lineId) %>%
  nrow()
  
valueBox(number_stops,
          icon = "fa-bus-alt")

    
})

```

### Distance

```{r}

renderValueBox({
 
distance <- data %>%
  filter(lineId == input$lineId) %>%
  filter(!is.na(Location_Easting)) %>%
  st_as_sf(coords = c("Location_Easting", "Location_Northing"), crs = 27700) %>%
  group_by(lineId) %>%
  summarise(do_union = FALSE) %>%
  st_cast("LINESTRING") %>%
  st_length()

distance <- (distance/1000) %>%
  round(digits = 2)
  
valueBox(paste(distance, "km"),
          icon = "fa-road")
})

```

### Time

```{r}

renderValueBox({
 
data_dist <- data %>%
  filter(lineId == input$lineId) %>%
  filter(!is.na(Location_Easting))

distance <- data_dist %>%
  st_as_sf(coords = c("Location_Easting", "Location_Northing"), crs = 27700) %>%
  group_by(lineId) %>%
  summarise(do_union = FALSE) %>%
  st_cast("LINESTRING") %>%
  st_length()

avg_speed <- (data_dist$Average_Speed[1] * 1.609)

time <- (((as.numeric(distance)/1000)*60)/(avg_speed)) %>%
  round(digits = 0)
  
valueBox(paste(time, "min"),
          icon = "fa-clock")
})


```

### Data provided by Transport for London: https://tfl.gov.uk/

Column {data-width=650}
-----------------------------------------------------------------------

### Route mapped for selected line

```{r}

# data <- st_as_sf(data, coords = c("Location_Easting", "Location_Northing"), crs = 27700)

bus_line <- reactive({
  data %>%
  filter(lineId == input$lineId) %>%
  filter(!is.na(Location_Easting)) %>%  
  st_as_sf(coords = c("Location_Easting", "Location_Northing"), crs = 27700) %>%
  select(lineId, id, name, lat, lon, Heading)  

 })

tmap_mode("view")

renderTmap({
  
net <- bus_line() %>%  
  group_by(lineId) %>%
  summarise(do_union = FALSE) %>%
  st_cast("LINESTRING") %>%
  as_sfnetwork(directed = TRUE)  
  
# now lets add points
tm_shape(st_as_sf(net, "edges")) + 
  tm_lines(id = c("Line" = "lineId"), 
           scale = 2) +
  tm_shape(st_as_sf(net, "nodes")) + 
    tm_dots(col = "#0019a8", 
            size = 0.2,
            id = "name") +
    tm_shape(bus_line()) + 
      tm_dots(col = "#dc241f",
              size = 0.04, 
              id = "name")  
 
})

```

Day and Night Network Comparison
===================================== 

Column {data-width=650}
-----------------------------------------------------------------------

### The day network is clearly more dense

```{r}
#edges_day <- read.csv("data/edges_day_start_end.csv") %>%
 # st_set_geometry(NULL)

net_day <- data_day %>%
  group_by(lineId) %>%
  filter(!is.na(Location_Easting)) %>%
  st_as_sf(coords = c("Location_Easting", "Location_Northing"), crs = 27700) %>%
  summarise(do_union = FALSE) %>%
  st_cast("LINESTRING") %>%
  mutate(length = as.numeric(st_length(geometry)))


tm_shape(net_day) + 
  tm_lines()

```

### TfL Bus lines serving London during daytime

```{r}

lines_serving_day <- nrow(net_day)

valueBox(lines_serving_day,
         icon = "fa-bus-alt",
         color = "red")

```

### Average length

```{r}

avg_length_day <- mean(net_day[["length"]]) / 1000
 
avg_stops_day <- data_day %>%
  group_by(lineId) %>%
  count(lineId) %>%
  ungroup() %>%
  summarise(Average_Stops = sum(n)/nrow(.)) %>%
  round(0)

valueBox(paste(avg_length_day %>% round(2), "km", "with", avg_stops_day, "stops on average"),
         icon = "fa-road")

```

### Average speed

```{r}

avg_speed_day <- mean(data_day[["Average_Speed"]])*1.609
 
valueBox(paste(avg_speed_day %>% round(2), "km/h"),
         icon = "fa-tachometer-alt")

```

Column {data-width=650}
-----------------------------------------------------------------------

### The night network's lines are longer on average

```{r}

net_night <- data_night %>%
  group_by(lineId) %>%
  filter(!is.na(Location_Easting)) %>%
  st_as_sf(coords = c("Location_Easting", "Location_Northing"), crs = 27700) %>%
  summarise(do_union = FALSE) %>%
  st_cast("LINESTRING") %>%
  mutate(length = as.numeric(st_length(geometry)))

tm_shape(net_night) + 
  tm_lines()

```

### TfL Bus lines serving London at night

```{r}

lines_serving_night <- nrow(net_night)

valueBox(lines_serving_night,
         icon = "fa-bus-alt",
         color = "red")

```

### Average length 

```{r}

avg_length_night <- mean(net_night[["length"]]) / 1000

avg_stops_night <- data_night %>%
  group_by(lineId) %>%
  count(lineId) %>%
  ungroup() %>%
  summarise(Average_Stops = sum(n)/nrow(.)) %>%
  round(0)

valueBox(paste(avg_length_night %>% round(2), "km", "with", avg_stops_night, "stops on average"),
         #subtitle = "Average number of stops",
         icon = "fa-road")
  
# stops_gauge <- 

```

### Average speed

```{r}

avg_speed_night <- mean(data_night[["Average_Speed"]])*1.609
 
valueBox(paste(avg_speed_night %>% round(2), "km/h"),
         icon = "fa-tachometer-alt")

```

Network Analysis - Nightbuses
===================================== 

Column {data-width=650}
-----------------------------------------------------------------------

### Visualization of shortest paths through nodes

```{r}

# simplified night bus network

bus_lines_mp <- read.csv("data/all_bus_lines_night.csv") %>%
  filter(!is.na(Location_Easting)) %>%
  group_by(parentId) %>%
  mutate(Location_Easting = mean(Location_Easting)) %>%
  mutate(Location_Northing = mean(Location_Northing))

# simplified bus stops

stop_mp <- bus_lines_mp %>%
  select(parentId, stationId, name, Location_Easting, Location_Northing) %>%
  unique()

# CREATE EDGE AND NODES FOR NETWORK DATA
# EVERY STATION FROM EVERY LINE GETS CONNECTED SEPERATELY

# with help of: https://www.r-spatial.org/r/2019/09/26/spatial-networks.html

# all nodes marked as start and end points, so that the data can be filtered by start and end

nodes <- bus_lines_mp %>%
  select(X, lineId,parentId, Location_Easting, Location_Northing, name) %>%
  rename("nodeID" = X) %>%
  mutate(start = "start", end = "end") %>%
  gather(start_end, end, start:end) #%>%

# first filter: start, then group per line, second filter: sort out the last stop per group as it can't be a start point

nodes_start <- nodes %>%
  filter(start_end == "start") %>%
  group_by(lineId) %>% 
  mutate(Index_Group = row_number()) %>%
  filter(Index_Group != max(n())) %>%
  ungroup()

# first filter: end, then group per line, second filter: sort out first entry per group as it can't be an end point

nodes_end <- nodes %>%
  filter(start_end == "end") %>%
  group_by(lineId) %>% 
  mutate(Index_Group = row_number()) %>%
  filter(Index_Group != 1) %>%
  ungroup() %>%
  mutate(nodeID = nodeID + 10000)  

# bind the two sets of start nodes and end nodes to create edges between each stop for all lines
    
edges <- cbind(D1 = nodes_start, D2 = nodes_end) %>%
  rename(c(D1.lineId, D1.Location_Easting, D1.Location_Northing, D2.Location_Easting, D2.Location_Northing),
         "lineId" = D1.lineId,
         "lat_start" = D1.Location_Easting, 
         "lon_start" = D1.Location_Northing, 
         "lat_end" = D2.Location_Easting, 
         "lon_end" = D2.Location_Northing) %>%
  select(lineId, D1.name, lat_start, lon_start, lat_end, lon_end) %>%
  unite(start_point, lat_start:lon_start, sep = "__") %>%
  unite(end_point, lat_end:lon_end, sep = "__") %>%
  mutate(Index = row_number()) %>% 
  gather(D1.name, point, start_point:end_point) %>%
  arrange(Index) %>%
  separate(point, c("lat", "lon"), sep = "__") %>%
  st_as_sf(coords = c("lat", "lon"), crs = 27700) %>%
  group_by(Index) %>%
  summarise(do_union = FALSE) %>%
  st_cast("LINESTRING")

# index edges with an ID
  
edges <- edges %>%
  mutate(edgeID = c(1:n()))

# actually not quite sure what is done in this part

nodes <- edges %>%
  st_coordinates() %>%
  as_tibble() %>%
  rename(edgeID = L1) %>%
  group_by(edgeID) %>%
  slice(c(1, n())) %>%
  ungroup() %>%
  mutate(start_end = rep(c('start', 'end'), times = n()/2)) %>%
  mutate(xy = paste(.$X, .$Y)) %>% 
  mutate(nodeID = group_indices(., factor(xy, levels = unique(xy)))) %>%
  select(-xy)

# next three chunks: pull the nodeIDs and create from-to relations within the edges data

source_nodes <- nodes %>%
  filter(start_end == "start") %>%
  pull(nodeID)

target_nodes <- nodes %>%
  filter(start_end == "end") %>%
  pull(nodeID)
  
edges = edges %>%
  mutate(from = source_nodes, to = target_nodes)

# delete non-unique entries

nodes <- nodes %>%
  distinct(nodeID, .keep_all = TRUE) %>%
  st_as_sf(coords = c('X', 'Y'), crs = 27700) %>%
  select(nodeID, geometry)

# combine edges and nodes to a complete network data

network <- sfnetwork(nodes, edges, directed = TRUE) %>%
  activate(edges) %>%
  mutate(length = st_length(geometry))
  
network <- network %>%  
  activate(nodes) %>%
  mutate(degree = centrality_degree()) %>%
  mutate(betweenness = centrality_betweenness(weights = as.numeric(length) + 0.0001)) %>%
  mutate(eigen = centrality_eigen()) %>%
  activate(edges) %>%
  mutate(betweenness = centrality_edge_betweenness(weights = as.numeric(length) + 0.0001)) # weights = as.numeric(length) + 0.0001

tm_shape(network %>% activate(edges) %>% as_tibble() %>% st_as_sf()) + 
  tm_lines() + 
  tm_shape(network %>% activate(nodes) %>% as_tibble() %>% st_as_sf()) +
  tm_dots(col = "betweenness",
          alpha = 0.6,
          size = "betweenness", 
          palette = "cividis")

```


Column {data-width=650}
-----------------------------------------------------------------------

### Betweenness centrality of edges and degree of nodes

```{r}

tm_shape(network %>% activate(edges) %>% as_tibble() %>% st_as_sf()) + 
  tm_lines() + 
  tm_shape(network %>% activate(nodes) %>% as_tibble() %>% st_as_sf()) +
  tm_dots(col = "degree",
          alpha = 0.6,
          size = "eigen", 
          palette = "cividis")

```

### Fifteen most linked bus stops 

```{r}

bus_stops_night_simple <- read.csv("data/all_bus_stops_night_simple.csv") %>%
  filter(!is.na(Location_Easting)) %>%
  st_as_sf(coords = c("Location_Easting", "Location_Northing"), crs = 27700)

nodes_degree <- network %>%
  activate(nodes) %>%
  as_tibble() %>%
  arrange(desc(degree)) %>%
  st_join(., bus_stops_night_simple) %>%
  st_set_geometry(NULL) %>%
  mutate(name = str_replace_all(name, c("Trafalgar Sq / Charing Cross Stn" = "Trafalgar Square",
                                        "Trafalgar Square / Charing Cross Stn" = "Trafalgar Square",
                                        "Northumberland Avenue / Trafalgar Square" = "Trafalgar Square"
                                        )))

nodes_degree_top <- nodes_degree %>%
  select(name, degree) %>%
  unique %>%
  slice(1:15)

renderPlot({
  
  ggplot(nodes_degree_top, aes(x = fct_reorder(name, degree), y = degree)) +
    geom_col(aes(fill = degree)) +
    scale_fill_viridis_b(alpha = 0.7) +
    coord_flip() +
    xlab("") +
    ylab("Degree") +
    theme_minimal()
  
})

  #select(nodeID, name, degree, lineId) %>%
  #group_by(name)
  #group_by(name)
#plot

#nodes_degree_plot <- plot_ly(nodes_degree_top, 
#                             x = ~degree, 
#                             y = ~reorder(name, degree),
#                             type = "bar",
#                             name = "TEST",
#                             orientation = "h",
#                             xaxis = list(title = ""))

#nodes_degree_plot <- nodes_degree_plot %>% 
#  layout(yaxis = list(showgrid = FALSE, showline = FALSE, showticklabels = TRUE, domain= c(0, 0.85)),
#         xaxis = list(zeroline = FALSE, showline = FALSE, showticklabels = TRUE, showgrid = TRUE))

#nodes_degree_plot

```
